\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}


\graphicspath {}


\title{Networks and Random Processes Assignment 2}
\author{Charlie Pilgrim - 1864704}
\date{October 2018}

\begin{document}

\maketitle


\section{Kingman's Coalescent}

\subsection{A}

$N_t$ is the number of particles at time t with $N_0=L$. The process $(N_t : t \geq 0)$ has the state space $\{1,...,L\}$

\subsubsection{Transition Rate of the process}

$r(n,n-1) = {L\choose 2} \, , \, n \geq 2$ 

QUESTION - WHAT ABOUT SAME STATE?
$r(n,n) = $

QUESTION - WHAT ABOUT OTHER STATES - HOW TO WRITE IT?
$r(n, y) = \, , \, y \neq n,n-1$

\subsubsection{Generator}

This is a jump process, so the generator is

$(\mathcal{L}f)(x) = \int_{\Re} r(x,y)[f(y)-f(x)]dy$

\bigskip

For this process

$(\mathcal{L}f)(n) = r(n,n-1)(f(n-1)-f(n))$

$(\mathcal{L}f)(n) = {n\choose 2} (f(n-1)-f(n))$

\subsubsection{Master Equation}

The master equation is

$\frac{d}{dt} \pi_t(n) = \pi_t(n+1)r(n+1,n) - \pi_t(n)r(n,n-1)$

$\frac{d}{dt} \pi_t(n) = \pi_t(n+1){n+1\choose 2} - \pi_t(n) {n \choose 2}$

QUESTION - IS THIS RIGHT?
QUESTION - IS THE NOTATION OKAY?
QUESTION - WHAT ABOUT EDGES? 


\subsubsection{Ergodicity}

The process is ergodic.

\subsubsection{Absorbing States}

The unique absorbing state is $N = 1$.

\subsubsection{Stationary Distributions}

Let a distribution $\pi = [N=1, N=2,... ,N=L]$

The unique stationary distribution is 

$\pi_0 = [1,0,...,0]$


\subsection{B - Mean Time to Asorption}

Show it is equal to $E(T) = 2(1-\frac{1}{L})$

Stefan said - sum expectation time across all steps( time to move to next state). Should be okay. 

Waiting time for each state is 

$\beta_n = \dfrac{1}{{n\choose 2}}$

The waiting times of differnet states are independent, so the mean time to absoprtion is

$E(T) = \sum_{i=1}^L \beta_i$

$E(T) = \sum_{i=1}^L \dfrac{1}{{i\choose 2}}$

$E(T) = \sum_{i=1}^L \dfrac{2}{i!}$

$E(T) = \dfrac{2L}{\sum_{i=1}^Li!}$

QUESTION - WTF? SUM OF FACTORIAL SEEMS WRONG


\subsection{C - Rescaled process}

Rescale the process to $N_t/L$

$(\mathcal{L}^Lf)(n/L) = \frac{1}{L} {n\choose2} (f(\frac{n-1}{L}) - f(\frac{n}{L}))$






\subsection{D}

\section{Ornstein-Uhlenbeck process}

\subsection{A}

\subsection{B}

\subsection{C}

\section{Moran Model and Wright-Fisher diffusion}

\subsection{A}

\subsection{B}

\subsection{C}

\subsection{D}

\subsection{E}

\subsection{F}



\end{document}





